---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data
1. Load the data (i.e. `read.csv()`)
```{r readdata, echo=TRUE}
unzip("activity.zip", exdir = "./data")
data <- read.csv("./data/activity.csv", sep=",", na.strings = "NA")
```
2. Process/transform the data (if necessary) into a format suitable for your analysis
```{r convertDate,echo=TRUE}
# Convert the column date to a date format
data$date <- as.Date(as.character(data$date), "%Y-%m-%d")
```

## What is mean total number of steps taken per day?
For this part of the assignment, you can ignore the missing values in the dataset.  
1. Make a histogram of the total number of steps taken each day
```{r totalStepsDay, echo=TRUE}
tot.steps.day <- tapply(data$steps, data$date, sum)
hist(tot.steps.day, main = "Histogram of Total Steps by Day", xlab = "Steps by Day")
```

2. Calculate and report the **mean** and **median** total number of steps taken per day
```{r mean_median, echo=TRUE}
mean(tot.steps.day, na.rm = TRUE)
median(tot.steps.day,na.rm = TRUE)
```

## What is the average daily activity pattern?
1. Make a time series plot (i.e. `type = "l"`) of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)
```{r, echo=TRUE}
# Load needed library
library(dplyr)

data_tbl <- tbl_df(data)
avg.steps.by.interval <- 
        data_tbl %>%
        group_by(interval) %>%
        summarize(avg_steps = mean(steps, na.rm = TRUE))

plot(avg.steps.by.interval, type="l", main = "Average dayly activity", xlab = "Interval", ylab = "Average steps")
```

2. Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?
```{r avg_max_steps_by_interval, echo=TRUE}
max.steps <- max(avg.steps.by.interval$avg_steps)
filter.max.steps <- filter(avg.steps.by.interval, avg_steps == max.steps)
interval.with.max.steps <- filter.max.steps$interval
interval.with.max.steps
```
The above calculation shows that interval ``r interval.with.max.steps`` contains the maximun number of steps (`r round(max.steps)` steps)

## Imputing missing values



## Are there differences in activity patterns between weekdays and weekends?
